Loop-Aware Optimizations in PyPy's Tracing JIT

### 2011-06-23 talk/iwtc11/paper.tex
sobel numbers
explain specialization

XXX explain that this is effectively type-specializing a loop

@@

XXX we either need to explain that we use C++ or consistently use C
\begin{figure}
\begin{center}
{\smaller
\begin{tabular}{|l|r|r|r|r|r|r|}
\hline
 & CPython & Psyco & PyPy  & PyPy & GCC \\
 &         &       & no LP &      & -O3 \\
\hline
conv3(1e5) & 77.89 & 9.52 & 1.77  & 0.68 &  0.59 \\
\hline
conv3(1e6) & 77.15 & 9.58 & 1.69  & 0.77 &  0.74 \\
\hline
conv3x3(1000) & 23.72 & 12.77 & 0.07  & 0.05 &  0.25 \\
\hline
conv3x3(3) & 23.85 & 12.77 & 0.10  & 0.07  &  0.27 \\
\hline
conv5(1e5) & 122.54 & 16.67 & 1.86  & 1.05 &  0.65\\
\hline
conv5(1e6) & 125.77 & 16.80 & 1.92 & 1.09 &  0.80 \\
\hline
dilate3x3(1000) & 23.29 & 12.99 & 0.41  & 0.39 &  0.26 \\
\hline
sobel(1000) & - & - & - & - & 0.20 \\
\hline
sqrt(Fix16) & 744.35 & 421.65 & 3.93  & 2.14  & 0.96 \\
\hline
sqrt(float) & 24.21 & 5.52 & 1.36 & 1.00 & 0.98\\
\hline
sqrt(int) & 20.84 & 1.78 & 2.26 & 1.82  & 0.80 \\
\hline
\hline
Variations & - & - & $\pm 0.03$ & $\pm 0.01$ & $\pm 0.01$ \\
\hline
\end{tabular}
}
\end{center}
\label{fig:benchmarks}
\caption{Benchmark Results in Seconds. Arrays of length $10^5$ and
  $10^6$ and matrixes of size $1000\times 1000$ and $1000000 \times
  3$ are used. This is indicated in the leftmost column. For the
  matrixes, only the number of rows are listed.} 
\end{figure}

>>>>

If all the optimizations presented above are applied, the resulting
optimized peeled loop will consist of a single integer addition
only. That is it will become type-specialized to the types of the
variables \lstinline{step} and \lstinline{y}, and the overhead of
using boxed values is removed.

@@

XXX we either need to explain that we use C++ or consistently use C
\begin{figure}
\begin{center}
{\smaller
\begin{tabular}{|l|r|r|r|r|r|r|}
\hline
 & CPython & Psyco & PyPy  & PyPy & GCC \\
 &         &       & no LP &      & -O3 \\
\hline
conv3(1e5) & 77.89 & 9.52 & 1.77  & 0.68 &  0.59 \\
\hline
conv3(1e6) & 77.15 & 9.58 & 1.69  & 0.77 &  0.74 \\
\hline
conv3x3(1000) & 23.72 & 12.77 & 0.07  & 0.05 &  0.25 \\
\hline
conv3x3(3) & 23.85 & 12.77 & 0.10  & 0.07  &  0.27 \\
\hline
conv5(1e5) & 122.54 & 16.67 & 1.86  & 1.05 &  0.65\\
\hline
conv5(1e6) & 125.77 & 16.80 & 1.92 & 1.09 &  0.80 \\
\hline
dilate3x3(1000) & 23.29 & 12.99 & 0.41  & 0.39 &  0.26 \\
\hline
sobel(1000) & 181.49 & 95.05 & 0.71 & 0.42 & 0.20 \\
\hline
sqrt(Fix16) & 744.35 & 421.65 & 3.93  & 2.14  & 0.96 \\
\hline
sqrt(float) & 24.21 & 5.52 & 1.36 & 1.00 & 0.98\\
\hline
sqrt(int) & 20.84 & 1.78 & 2.26 & 1.82  & 0.80 \\
\hline
\hline
Variations & - & - & $\pm 0.03$ & $\pm 0.01$ & $\pm 0.01$ \\
\hline
\end{tabular}
}
\end{center}
\label{fig:benchmarks}
\caption{Benchmark Results in Seconds. Arrays of length $10^5$ and
  $10^6$ and matrixes of size $1000\times 1000$ and $1000000 \times
  3$ are used. The one used in each benchmark is indicated in
  the leftmost column. For the matrixes, only the number of rows are
  specified.} 
\end{figure}
